---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CRTConjoint

<!-- badges: start -->
<!-- badges: end -->

The goal of CRTConjoint is to use the conditional randomization test (CRT) to 
test for various hypothesis in conjoint experiments. In particular, `CRT_pval`
aims to test whether a factor matters in any way. For example, does education matter
in immigration preferences given other attributes of the candidate.

## Installation

You can install CRTConjoint from [GitHub](https://github.com/daewoongham97/CRTConjoint) with:
``` r
# install.packages("devtools")
devtools::install_github("daewoongham97/CRTConjoint")
```

or directly from CRAN with:
``` r
install.packages("CRTConjoint")
```

## Example

This is a basic example which shows you how to test whether education matters
for immigration preferences.

```{r example, eval = FALSE}
library(CRTConjoint)

#
data("immigrationdata")
form = formula("Y ~ FeatEd + FeatGender + FeatCountry + FeatReason + FeatJob +
FeatExp + FeatPlans + FeatTrips + FeatLang + ppage + ppeducat + ppethm + ppgender")
left = colnames(immigrationdata)[1:9]
right = colnames(immigrationdata)[10:18]

## Not run: 
# Testing whether edcuation matters for immigration preferences
education_test = CRT_pval(formula = form, data = immigrationdata, X = "FeatEd",
 left = left, right = right, non_factor = "ppage", B = 100, analysis = 2)
education_test$p_val

```

