---
title: "Using CRTConjoint"
output: rmarkdown::html_vignette
description: >
  This vignette describes the basics of how to use CRTConjoint package along with
  how to effectively use this in the Amazon Web Services parallel computing cluster.
vignette: >
  %\VignetteIndexEntry{Using CRTConjoint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CRTConjoint)
```

This vignette will go through examples cases listed in the package for all the main 
functions with details on how to use the inputs. Finally, the last section contains
detailed steps on how to use CRTConjoint with the Amazon Web Services computing cluster.

## Understanding CRT_pval
To begin, we start with understanding the main function `CRT_pval`. All examples 
use the immigration choice conjoint experiment data from Hainmueller et. al. (2014).
We first load this data from the package.
```{r}
data("immigrationdata")
```
Each row consists a pair of immigrant candidates that were shown to respondents.
For example, the first row shows that the left immigrant candidate was a Male from Iraq
who had a high school degree while the right immigrant candidate was a Female from France
with no formal education. The respondent who evaluated this task was a 20 year old 
college educated White Male, who voted for the left immigrant candidate. 

In the first example, we aim to understand if the candidate's education matters
for immigration preferences. To test this, `CRT_pval` requires users to specify
all factors and respondent factors of interest in the `formula` input along with 
the binary response. `CRT_pval` requires us to additionally specify which of 
the factors are the left and right profiles. 
```{r}
form = formula("Y ~ FeatEd + FeatGender + FeatCountry + FeatReason + FeatJob +
FeatExp + FeatPlans + FeatTrips + FeatLang + ppage + ppeducat + ppethm + ppgender")
left = colnames(immigrationdata)[1:9]
right = colnames(immigrationdata)[10:18]
left; right
```
Users can see that the left and right profile factors are aligned, i.e., the first entry
is the education for both the left and right profiles. It is important that they
are expected to be aligned. We also note that the formula only contains the factors
for the left profile. This is sufficient as the algorithm will take the left and right 
input to use both left and right profile attributes for testing the hypothesis. Lastly,
we include all respondent characteristics (ppage, ppeducat, etc.) to boost power. 
We are ready to now run `CRT_pval` to test whether education matters.
```{r, eval = FALSE}
education_test = CRT_pval(formula = form, data = immigrationdata, X = "FeatEd",
 left = left, right = right, non_factor = "ppage", B = 100, analysis = 2)
education_test$p_val
```
We again note `X = "FeatEd"` is sufficient to clarify which factor we are testing for. 
Because the function assumes all attributes in conjoint experiments are of class factor,
if there are variables that are not factor class, for example
respondent age (ppage), the function must know which are these non-factor variables. 
Furthermore, to save time we only run it for $B = 100$ resamples. Lastly, 
we set $analysis = 2$ to also allow the function to spit out two of the strongest 
interactions that contributed to the observed test statistic. We note that this 
is purely for exploratory purposes. 

The output should contain not only the $p$-value but also the observed test statistic, 
all the resampled test statistic, etc. The function will also show a progress bar 
to show the percentage of resamples the user is finished with. 

### Using constrained randomization design
Since the immigrant's education was uniformly sampled, we did not need to specify 
the design because the default design was uniform. However, there are some factors
that used more complex design. One such factor was job (FeatJob). For this experiment,
the candidate's occupation could only be financial analyst, computer programmer, research
scientist, and doctor if their education degree was equivalent to at least some level of 
college. Because this constrained uniform design is popular in conjoint experiments,
we allow the function to account for this design so long as the user specifies the 
constraints. We now show, using the same example, how to specify this constraint. 
```{r, eval = FALSE}
constraint_randomization = list() # (Job has dependent randomization scheme)
constraint_randomization[["FeatJob"]] = c("Financial analyst","Computer programmer",
"Research scientist","Doctor")
constraint_randomization[["FeatEd"]] = c("Equivalent to completing two years of
college in the US", "Equivalent to completing a graduate degree in the US",
 "Equivalent to completing a college degree in the US")
```
The `constraint_randomization` list is a list of length two. The first element contains
the levels of job that can only be randomized with certain levels of education.
The second element of the list contains the levels of education that allows the aforementioned
jobs to have positive probability. The listed levels are assumed to match the levels in the supplied 
data. Additionally, the names of the list is also assumed to match the column names in the supplied data.
We note that the user only has to supply the constraint for either the left or right factor
and the function will assume the constraint randomization scheme is the same for both left 
and right factors, i.e., same for FeatJob_2 and FeatEd_2. 
```{r, eval = FALSE}
job_test = CRT_pval(formula = form, data = immigrationdata, X = "FeatJob",
left = left, right = right, design = "Constrained Uniform",
constraint_randomization = constraint_randomization, non_factor = "ppage", B = 100)
job_test$p_val
```
Once, we have the constraint list, we input it into the `constraint_randomization` input, 
after stating that the `design = "Constrained Uniform"`. We supply other examples
when a user has a nonuniform (but no constraints) design and how to force a variable
to include as an interaction in the examples. 

## Understanding extensions of CRT
We provide three other CRT functions that take similar inputs but tests regularity conditions
often invoked in conjoint experiments. The first aims to test the profile order 
effect (`CRT_profileordereffect`), i.e., whether or not being in the left or right has 
any impact on the response. The syntax for such a test is straightforward.
```{r, eval = FALSE}
profileorder_test = CRT_profileordereffect(formula = form, data = immigrationdata,
 left = left, right = right, B = 100)
profileorder_test$p_val
```
Testing the carryover effect and fatigue effect is slightly more complex. When testing
the carryover effect, it requires resampling all the left and right factors. The default
`design = "Uniform"` assumes that all factors were uniformly sampled. If that is not the case,
(which the immigration example is not) we need to supply our own resamples. To do this,
we build our resampling function,
```{r, eval = FALSE}
resample_func_immigration = function(x, seed = sample(c(0, 1000), size = 1), left_idx, right_idx) {
 set.seed(seed)
 df = x[, c(left_idx, right_idx)]
 variable = colnames(x)[c(left_idx, right_idx)]
 len = length(variable)
 resampled = list()
 n = nrow(df)
 for (i in 1:len) {
   var = df[, variable[i]]
   lev = levels(var)
   resampled[[i]] = factor(sample(lev, size = n, replace = TRUE))
 }

 resampled_df = data.frame(resampled[[1]])
 for (i in 2:len) {
   resampled_df = cbind(resampled_df, resampled[[i]])
 }
 colnames(resampled_df) = colnames(df)

 #escape persecution was dependently randomized
 country_1 = resampled_df[, "FeatCountry"]
 country_2 = resampled_df[, "FeatCountry_2"]
 i_1 = which((country_1 == "Iraq" | country_1 == "Sudan" | country_1 == "Somalia"))
 i_2 = which((country_2 == "Iraq" | country_2 == "Sudan" | country_2 == "Somalia"))

 reason_1 = resampled_df[, "FeatReason"]
 reason_2 = resampled_df[, "FeatReason_2"]
 levs = levels(reason_1)
 r_levs = levs[c(2,3)]

 reason_1 = sample(r_levs, size = n, replace = TRUE)

 reason_1[i_1] = sample(levs, size = length(i_1), replace = TRUE)

 reason_2 = sample(r_levs, size = n, replace = TRUE)

 reason_2[i_2] = sample(levs, size = length(i_2), replace = TRUE)

 resampled_df[, "FeatReason"] = reason_1
 resampled_df[, "FeatReason_2"] = reason_2

 #profession high skill fix
 educ_1 = resampled_df[, "FeatEd"]
 educ_2 = resampled_df[, "FeatEd_2"]
 i_1 = which((educ_1 == "Equivalent to completing two years of college in the US" |
  educ_1 == "Equivalent to completing a college degree in the US" |
  educ_1 == "Equivalent to completing a graduate degree in the US"))
 i_2 = which((educ_2 == "Equivalent to completing two years of college in the US" |
 educ_2 == "Equivalent to completing a college degree in the US" |
 educ_2 == "Equivalent to completing a graduate degree in the US"))


 job_1 = resampled_df[, "FeatJob"]
 job_2 = resampled_df[, "FeatJob_2"]
 levs = levels(job_1)
 # take out computer programmer, doctor, financial analyst, and research scientist
 r_levs = levs[-c(2,4,5, 9)]

 job_1 = sample(r_levs, size = n, replace = TRUE)

 job_1[i_1] = sample(levs, size = length(i_1), replace = TRUE)

 job_2 = sample(r_levs, size = n, replace = TRUE)

 job_2[i_2] = sample(levs, size = length(i_2), replace = TRUE)

 resampled_df[, "FeatJob"] = job_1
 resampled_df[, "FeatJob_2"] = job_2

 resampled_df[colnames(resampled_df)] = lapply(resampled_df[colnames(resampled_df)], factor )

 return(resampled_df)
}
```
This resampling function takes the data (x) as an input and given the indexes of the
left and right profile attributes, it returns a completely new resampled dataframe of 
the same dimension as x. As stated above, because testing the carryover effect requires
resampling all the attributes B times, we must supply all the manually 
resampled dataframes. To do this we store them in a list of length B, each containing
a resampled data of all the left and right attributes. We store this in `own_resamples`.
```{r, eval = FALSE}
carryover_df = immigrationdata
own_resamples = list()
B = 100
for (i in 1:B) {
 newdf = resample_func_immigration(carryover_df, left_idx = 1:9, right_idx = 10:18, seed = i)
 own_resamples[[i]] = newdf
}
```
Lastly, the carryover test requires a column that indicates the task evaluation number
for each row. In the immigration experiment, each respondent rated five tasks and the data
is sorted by each task, i.e., the first five rows are the first five tasks for the first respondent. 
Consequently, we can define a new column, task, that iterates 1 to 5 and run the main function.
NOTE: it is important that the task variable has no missing tasks, i.e., a respondent that
only rated four tasks while all other respondents rated five tasks. If there is such a respondent,
please drop it before using this function. 

```{r, eval = FALSE}
J = 5
carryover_df$task = rep(1:J, nrow(carryover_df)/J)

carryover_test = CRT_carryovereffect(formula = form, data = carryover_df, left = left,
right = right, task = "task", supplyown_resamples = own_resamples, B = B)
carryover_test$p_val
```

Lastly, to test the fatigue effect, we only need to additionally specify the respondent
index. Like the task column, we similarly repeat 1 to 200 five times and store it
in a new column called respondent. 
```{r, eval = FALSE}
fatigue_df = immigrationdata
fatigue_df$task = rep(1:J, nrow(fatigue_df)/J)
fatigue_df$respondent = rep(1:(nrow(fatigue_df)/J), each = J)

fatigue_test = CRT_fatigueeffect(formula = form, data = fatigue_df, left = left,
right = right, task = "task", respondent = "respondent", B = 100)
fatigue_test$p_val
```


## Using CRTConjoint with Amazon Web Services
An optional argument for all CRT functions in this package is the `num_cores` input.
Currently it is set at 2 cores as the default. Although 2 cores may be fine for someone
who is doing exploratory work with $B = 100$, for final reported $p$-values a much higher
$B$ is recommended. A typical Mac laptop will only support up to `num_cores = 4`, which
may be unsatisfactory for researchers. For researchers with their own computing cluster, 
this section is not applicable for them. However, for those with no easy access to computing
cluster, we write this section to easily use our functions in powerful computing clusters
provided by Amazon Web Services (AWS). 

We will leverage the RStudio Server in AWS maintained and provided by:
https://www.louisaslett.com/RStudio_AMI/. We will now list the steps needed to use AWS Rstudio.

* Step 1) Sign up for an AWS account. Then login.
* Step 2) Go to the above link: https://www.louisaslett.com/RStudio_AMI/ and click
on the region closest to user's local region. For example, we click on US East, Ohio
ami-07038... 
* Step 3) This will redirect you to a new page. Under instance type select the desired
instance type. The number next to vCPU indicates the number of cores available. We
recommend c5a.16xlarge with 64 cores. Please note that this costs 2.4 USD per hour.
The user will be billed per hour. There are cheaper options. 
* Step 4) After selecting instance type, scroll down to "Network settings". Make sure
"Allow SSH Traffic" is set to "Anywhere". Also tick "Allow HTTP traffic from the internet". 
* Step 5) Press the "launch instance" orange button.
* Step 6) After being redirected, MANUALLY copy and paste the link below "Public 
IPv4 DNS". Then paste it on a new tab on google chrome. DO NOT press the open address
button that is there as it does not work. 
* Step 7) You will be prompted (might take a minute) to a RStudio login page. 
Username: rstudio and password is your instance ID.
* Step 8) You can now enjoy RStudio on this AWS server. Install CRTConjoint package
set num_cores to the desired cores and run the CRT functions faster! 

We recommend that the user has already written all necessary code to run on the AWS. 
Since users will be charged hourly, it is recommended that they can directly run the code. 
If the user runs with 50 cores, the runtime for a typical conjoint experiment should not 
exceed more than 5-10 minutes for a single $p$-value with $B = 2000$,
thus we do not believe any user to have to incur much fees. For any questions please 
do not hesitate to contact me at: daewoongham at g dot harvard dot edu










